---
title: "Data exploration of Accidents"
author: "Céline Guex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

```{r loading-data}
load("Data/Accidents.RData")
```

# A first exploration of the dataset

```{r}
head(Accidents)
```

```{r visu-y}
ggplot(Accidents, aes(Acc))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1)+
  ggtitle("Density of accidents")
```

```{r visu2}

ggplot(Accidents, aes(x=Traffic,y=Acc))+
  geom_point(shape=1)+
  ggtitle("Accidents vs traffic")

ggplot(Accidents, aes(Acc))+
  geom_bar(alpha =0.5)+
  facet_grid(rows=vars(Limit), space = "free" )+
  ggtitle("Count of accidents, by limit of speed")

ggplot(Accidents, aes(x=Acc))+
  geom_bar(shape=1)+
  facet_grid(~Type)+
  ggtitle("Accidents by type of tunnel")

ggplot(Accidents, aes(x=Acc, fill=Type))+
  geom_bar(position="fill")+
  ggtitle("Repartition of Accidents by tunnel type")

ggplot(Accidents, aes(x=Type))+
  geom_bar()+
  ggtitle("Repartition of tunnels")

ggplot(Accidents, aes(x=HGV,y=Traffic))+
  geom_point()+
  ggtitle("Traffic VS HGV")

ggplot(Accidents, aes(x=Acc, fill=Urban))+
  geom_bar(position = "dodge")+
  ggtitle("Accidents VS Urban ")

ggplot(Accidents, aes(x=Slope, y=Acc))+
  geom_point(shape=1)+
  ggtitle("Accidents VS slope")

ggplot(Accidents, aes(x=Acc, fill=Company))+
  geom_bar(position="fill")+
  ggtitle("Accidents by compagny")

ggplot(Accidents, aes(x=Tunnel, y=Acc))+
  geom_point()+
  ggtitle("Accidents by Tunnel")

ggplot(Accidents, aes(x=Acc, fill=SlopeType))+
  geom_bar(position = position_dodge(0.5))+
  ggtitle("Accidents by SlopeType")

ggplot(Accidents, aes(x=Acc, fill=SlopeType))+
  geom_bar(position = "fill")+
  ggtitle("Accidents by SlopeType")

ggplot(Accidents, aes(x=Traffic, y=Acc, shape=SlopeType))+
  geom_point()+
  facet_grid(~SlopeType)
  ggtitle("Accidents VS Traffic, by Slope type")
  
ggplot(Accidents, aes(x= Slope,y=Acc))+
  geom_bin_2d()+
  ggtitle("Accidents VS Slope")
```

## First linear model

Without informations on tunnels or companies !
```{r linear_model}
lm1 <- lm(Acc~Traffic+HGV+Slope+Urban+Type+Length+Limit+SlopeType+Width+Lanes, data=Accidents)
summary(lm1)
```
From this first linear regression, it seems that "(Intercept)", Traffic", "Type", "Urban", "Length", "Limit", "Slope" and "Width" are significant. However, the Adjusted R-squared is very small, which may indicate that this model is not appropriate. <br>
Let us now look what problems we might have :
```{r check_lm1_1}
#Checking linearity
y<-fitted.values(lm1)
r<-rstandard(lm1)
ggplot(data.frame(y,r),aes(y,r))+
  geom_point(shape=1)+
  # geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
  ggtitle("standardized residuals vs fitted values")
```

Here, it seems that there is a pattern appearing when few accidents occurs, thus independence of the residuals $r$ and the fitted values $\hat{y}$ might fail.

```{r check_lm1_2}
#checking variance
r2 <- abs(r)
ggplot(data.frame(y,r2),aes(y,r2))+
  geom_point(shape=1)+
  # geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
  ggtitle("absolute standardized residuals vs fitted values")
```

Here, the pattern suggest that $var(r)$ might increase with the mean increasing. Not good either...

```{r check_lm1_3}
# checking independance
acf(r, lag.max = NULL,
    type = c("correlation", "covariance", "partial"),
    plot = TRUE, na.action = na.fail, demean = TRUE)
pacf(r, lag.max = NULL,
     plot = TRUE, na.action = na.fail, demean = TRUE)
```
It seems that there are no dependances within the standardized residuals... (I am not an expert with correlograms)

```{r check_lm1_4}
#checking normality
ggplot(lm1, aes(sample=r))+
  stat_qq()+
  geom_abline(slope=1,intercept = 0)+
  ggtitle("Normal Q-Q plot")
```

Here is the normal Q-Q plot with a line of slope 1 and intercept 0. It seems that there are heavy tails in this case !

