---
title: "Data exploration of Fires"
author: "Céline Guex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(corrplot)
library(plotly)
library(ggExtra)
```

```{r loading-data}
load("Data/Fires.RData")
```

# A first exploration of the dataset

```{r header}
head(Fires)
#str(Fires)
```

```{r visu1}
# Fires[1:nrow(Fires),2:ncol(Fires)] %>% pivot_longer(everything()) %>%
#   ggplot(aes(value)) + 
#   facet_wrap(~ name, scales = "free",ncol=3) +
#   geom_histogram()+
#   ggtitle("First look on the data")

bw=0.3
ggplot(Fires, aes(x=Fires))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1)+
  ggtitle("Repartition of y")
# yp<-density(Fires$Fires)
# pp<-rpois(length(yp),0.8)

p<-data.frame(rexp(100,2))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=bw)+
  ggtitle("exp(2)")

p<-data.frame(rpois(10000,0.8))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=bw)+
  ggtitle("pois(0.8)")
  

```

```{r visu2}

ggplot(Fires, aes(x=Traffic,y=Fires))+
  geom_point(shape=1)+
  ggtitle("Fires vs traffic")

ggplot(Fires, aes(Fires, fill=as.factor(Limit)))+
  geom_bar(position = position_dodge(0.5), alpha =0.5)+
  ggtitle("Count of fire, by limit of speed")

ggplot(Fires, aes(x=Fires))+
  geom_bar(shape=1)+
  facet_grid(~Type)+
  ggtitle("Fires by type of tunnel")

ggplot(Fires, aes(x=Fires, fill=Type))+
  geom_bar(position="fill")+
  ggtitle("Repartition of fires by tunnel type")

ggplot(Fires, aes(x=Type))+
  geom_bar()+
  ggtitle("Repartition of tunnels")

ggplot(Fires, aes(x=HGV,y=Traffic))+
  geom_point()+
  ggtitle("Traffic VS HGV")

ggplot(Fires, aes(x=Fires, fill=Urban))+
  geom_bar(position = "dodge")+
  ggtitle("Fire VS Urban ")

ggplot(Fires, aes(x=Slope, y=Fires))+
  geom_point(shape=1)+
  # geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95,
  #             color="red")+
  ggtitle("Fires VS slope")

ggplot(Fires, aes(x=Fires, fill=Company))+
  geom_bar(position="fill")+
  ggtitle("Fires by compagny")

ggplot(Fires, aes(x=Tunnel, y=Fires))+
  geom_point()+
  ggtitle("Fires by Tunnel")

ggplot(Fires, aes(x=Fires, fill=SlopeType))+
  geom_bar(position = position_dodge(0.5))+
  ggtitle("Fires by SlopeType")

ggplot(Fires, aes(x=Fires, fill=SlopeType))+
  geom_bar(position = "fill")+
  ggtitle("Fires by SlopeType")

ggplot(Fires, aes(x=Traffic, y=Fires, shape=SlopeType))+
  geom_point()+
  facet_grid(~SlopeType)+
  ggtitle("Fires VS Traffic, by Slope type")
  
ggplot(Fires, aes(x= Slope,y=Fires))+
  geom_bin_2d()+
  ggtitle("Fires VS Slope")
```

## First linear model
Without informations on tunnels or companies !

```{r some_regressions}
# lm1 <-lm(Fires~.,data=Fires)
# summary(lm1)
lm1 <- lm(Fires~Traffic+HGV+Slope+Urban+Type+Length+Limit+SlopeType, data = Fires)
summary(lm1)
```
From this first linear regression, it seems that "Traffic", "HGV", "Urban" and "Length" are significant. However, the Adjusted R-squared is very small, which may indicate that this model is not appropriate. <br>
Let us now look what problems we might have :
```{r check_lm1_1}
#Checking linearity
y<-fitted.values(lm1)
r<-rstandard(lm1)
ggplot(data.frame(y,r),aes(y,r))+
  geom_point(shape=1)+
  geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
  ggtitle("standardized residuals vs fitted values")
```

Here, it seems that there is a pattern appearing, thus independence of the residuals $r$ and the fitted values $\hat{y}$ might fail.

```{r check_lm1_2}
#checking variance
r2 <- abs(r)
ggplot(data.frame(y,r2),aes(y,r2))+
  geom_point(shape=1)+
  geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
  ggtitle("absolute standardized residuals vs fitted values")
```

Here, the pattern suggest that $var(r)$ might increase with the mean increasing. Not good either...

```{r check_lm1_3}
# checking independance
acf(r, lag.max = NULL,
    type = c("correlation", "covariance", "partial"),
    plot = TRUE, na.action = na.fail, demean = TRUE)
pacf(r, lag.max = NULL,
     plot = TRUE, na.action = na.fail, demean = TRUE)
```
It seems that there are no dependances within the standardized residuals... (I am not an expert with correlograms)

```{r check_lm1_4}
#checking normality
ggplot(lm1, aes(sample=r))+
  stat_qq()+
  geom_abline(slope=1,intercept = 0)+
  ggtitle("Normal Q-Q plot")
```

Here is the normal Q-Q plot with a line of slope 1 and intercept 0. The result of normality is quite correct. Maybe a bit right-skewed.
We can notice that we have an outlier !!!
