---
title: "Project Regression Methods - 2022"
author: "Céline Guex - sciper n°310673, Candice Baud - sciper n°359523"
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Initial data analysis

## Fires

```{r explo-fires, file='Fires/Fires_explo.R'}
```

## Accidents

```{r explo-acc, file='Accidents/Accidents_explo.R'}
```

## Breakdowns

```{r explo-breakdowns, file='Breakdowns/Breakdowns_explo.R'}
```

# Model Fitting

## Fires

For the tables, see file "Fires/Fires_modelglm.tex", "Fires/Fires_modelglmr_fixed.tex" that are generated by "Fires/Fires_diagnostics.R"

### Poisson model

```{r model-fires1, file='Fires/Fires_models.R'}
```

### Poisson mixed 

```{r model-fires2, file='Fires/Fires_models2.R'}
```

## Accidents 

For the tables, see file "Accidents/Accidents_modelglm.tex", "Accidents/Accidents_modelglmr_fixed.tex" that are generated by "Accidents/Accidents_diagnostics.R"

### Poisson model

```{r model-acc1, file='Accidents/Accidents_models.R'}
```

### Poisson mixed model

```{r model-acc2, file='Accidents/Accidents_models2.R'}
```

## Breakdowns

For the table, see file "Breakdowns/Breakdowns_model.tex" that is generated by "Breakdowns/Breakdowns_diagnostics.R

```{r model-break, file='Breakdowns/Breakdowns_models.R'}
```



# Discussion

## Fires

```{r diag-fires, file='Fires/Fires_diagnostics.R'}
```


## Accidents

```{r diag-acc, file='Accidents/Accidents_diagnotics.R'}
```


## Breakdowns

```{r diag-break, file='Breakdowns/Breakdowns_diagnostics.R'}
```


# Conclusion