---
title: "Fires Candice"
author: "candice"
date: "2022-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r Packages, echo=FALSE, include = FALSE}
lapply(c("dplyr","chron","ggplot2","tidyr","questionr","survival","forcats",
         "data.table","table1","lubridate", "ggpubr","viridis","finalfit",
         "ggpubr", "ggthemes", "gridExtra", "tidyverse", "rstatix","ggsci",
         "wesanderson","kableExtra", "naniar","boot","scales","psych","ggcorrplot","ggradar", "caret", "gam"), 
       library, character.only=TRUE)

Sys.setlocale("LC_TIME", "English")
setwd("C:/Users/candi/Desktop/ETUDES/EPFL1A/semestre 1/regression methods/project")
load("C:/Users/candi/Desktop/ETUDES/EPFL1A/semestre 1/regression methods/project/Fires.RData")
```

## Exploratory
The distribution looks exponential or poisson, and ther might be an outlier after 10. 
```{r plot fires, echo=FALSE}
Fires %>% group_by(Fires) %>% summarise(number=n()) %>% mutate(freq = number/sum(number)) %>% 
  ggplot(aes(x=Fires, y = freq)) + geom_bar(stat='identity')
```

Who is the possible outlier ? 

```{r who is the possible outlier, echo=FALSE}
Fires %>% filter(Fires >10)
```


```{r correlation, echo=FALSE}
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}


pairs(~ Fires + Traffic + HGV + Slope + Limit + Length,data=Fires, upper.panel = panel.cor, diag.panel = panel.hist)
``` 

It looks like Fires are correlated with Traffic, HGV and Length. 
Length and HGV are correlated too so we should take this is account when we will try to fit the model.
```{r HGV and Length, echo = FALSE}
Reg1 <- lm(Fires$HGV ~ log(Fires$Length))
plot(log(Fires$Length), Fires$HGV)
abline(Reg1)
```

```{r visualisations, echo = FALSE}
plot(log(Fires$Traffic), Fires$Fires)

plot(Fires$HGV, Fires$Fires)

plot(Fires$Length, Fires$Fires) #exponential distribution ? 

```
On papers I have read, they mention that Fires and Slope are significatively correlated, let's see: 

```{r}
plot((Fires$Slope), Fires$Fires)

slope.lm <- lm(Fires~Slope, data=Fires)
summary(slope.lm)

```
The regression is clearly not significant as we might have expected from the graph. 


## Regressions attempts 
Traffic, HGV and Length seem to be significant.

```{r regression with all parameters, echo=FALSE}
Reg2 <- lm(Fires ~ Traffic + HGV + Slope + Urban + Type + Length + Limit + SlopeType, data=Fires)
summary(Reg2)
```
Doing the regression on only those 3 is in fact not that bad in terms of pvalue. 
```{r regression with Traffic HGV Length, echo=FALSE}
Reg3 <- lm(Fires ~ Traffic + HGV + Length, data=Fires)
summary(Reg3)
```

Let's look at the residuals. They are not normal at all (QQplot), there is no homoscedasticity, and we have some outliers.

```{r residuals of Reg3, echo=FALSE}
plot(Reg3)
```

```{r regression tests, echo=FALSE}
Reg4 <- lm(Fires ~ log(Traffic) + HGV + Length, data=Fires)
summary(Reg4)
plot(Reg4)

Reg5 <- lm(Fires ~ Traffic + HGV + log(Length), data=Fires)
summary(Reg5)
plot(Reg5)
#Reg4 is better

r<-rstandard(Reg5)
acf(r, lag.max = NULL,
    type = c("correlation", "covariance", "partial"),
    plot = TRUE, na.action = na.fail, demean = TRUE)
pacf(r, lag.max = NULL,
     plot = TRUE, na.action = na.fail, demean = TRUE)
```
## Econometric approach
We try to look at the effect of HGV in our model. First we regress HGV on Length, the residuals will then capture the pure effect of HGV (because the effect of Length on HGV is already expressed in the betaOLS). So then, I regress Fires on the residuals. The result is that HGV is not that significant. 
```{r Econometric approach}
Reg6 <- lm(HGV ~ Length, data = Fires)
summary(Reg6)

r <- residuals(Reg6)
plot(r)

#by regressing on the residual, we can eliminate the effect of HGV on Length and keep only the effect of Length
Reg7 <- lm(Fires ~r, data=Fires)
summary(Reg7)
plot(Reg7)
```

By doing it the other way around, we can capture the effect of Length on Fires. It appears to have a greater effect than HGV. However, it is still limited. The thing that appears clearly is the fact that HGV and Length have a strong correlation.

```{r Econometric approach 2}
Reg8 <- lm(Length ~ HGV, data = Fires)
summary(Reg8)

r <- residuals(Reg8)
plot(r)

Reg9 <- lm(Fires ~ r, data=Fires)
summary(Reg9)
plot(Reg9)
```

## Poisson model
First, I look at the mean to find the parameter.
```{r Poisson mean, echo =FALSE}
mean(Fires$Fires)
```

Then I plot the distribution and look at how it looks.
```{r Plot things}
p<-data.frame(rpois(10000,mean(Fires$Fires)))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=0.2)+
  ggtitle("Poisson model")
``` 

Trying to fit a Poisson model. The summary looks not so bad but the residuals are clearly not normal and there are some outliers. 
```{r Poisson model, echo =FALSE}
m1 <- glm(Fires ~ Length + HGV + Traffic, family="poisson", data=Fires)
summary(m1)
plot(m1) #the residuals look terribly bad

m1$aic
```
Second Poisson model : the summary is way better, I tried to plot the prediction but I am not sure I have understood what it does. Also the residuals look worse. The aic is slightly bigger. 
```{r}
m2<- glm(Fires ~ HGV + Traffic, family="poisson", data=Fires)
summary(m2) #the summary looks pretty cool
plot(m2) #residuals are still bad

m2$aic

Fires$hat <- predict(m2, type='response')
Fires <- Fires[with(Fires, order(Traffic, HGV)), ]
ggplot(Fires, aes(x = HGV, y = hat, colour = Traffic)) +
  geom_point(aes(y = Fires), alpha=.5, position=position_jitter(h=.2)) +
  geom_line(size = 1) +
  labs(x = "HGV", y = "Fires")
```

## Polynomial regression 
With HGV, the summary is good, but the residuals don't look very good when I plot them. 
```{r HGV}
glm.fit <- glm(Fires~poly(HGV, 3), data=Fires)
summary(glm.fit)
confint(glm.fit)

plot(fitted(glm.fit), residuals(glm.fit))
```
The summary for Traffic isn't as good as for HGV but the residuals look better.
```{r Traffic}
glm.fit2 <- glm(Fires~poly(Traffic, 3), data=Fires)
summary(glm.fit2)
confint(glm.fit2)

plot(fitted(glm.fit2), residuals(glm.fit2))
```


## Variable selection ? 
I think this part is not useful but anyway I let it.
```{r}
nullmodel <- lm(Fires ~ 1, data = Fires)
fullmodel <- lm(Fires ~ ., data = Fires)
```

Forward selection
```{r}
model.step.f <- step(nullmodel, scope = list(lower = nullmodel, upper = fullmodel), direction = 'forward')
``` 

Backward elimination
```{r}
model.step.b <- step(fullmodel, direction = 'backward')
``` 

Stepwise selection
```{r}
model.step <- step(nullmodel, scope = list(lower = nullmodel, upper = fullmodel), direction = 'both')
``` 


AICs and BICs
```{r}
AIC(model.step.f)
BIC(model.step.f)
AIC(model.step.b)
BIC(model.step.b)
AIC(model.step)
BIC(model.step)
```

## Trying other things that might be bad
```{r test with gam, echo=FALSE}
fit.gcv <- gam(Fires~Length-1+s(HGV),data=Fires)
summary(fit.gcv)
fit.plot.gcv <- plot(fit.gcv,seWithMean=TRUE,xlab="Length")
fit.gcv$aic
```

```{r test with gam 2, echo=FALSE}
fit.gcv <- gam(Fires~HGV-1+s(HGV),data=Fires)
summary(fit.gcv)
fit.plot.gcv <- plot(fit.gcv,seWithMean=TRUE,xlab="HGV")
fit.gcv$aic

```

```{r with random effects of tunnels}
#fit.gcv.re <- gam(Fires~s(Tunnel,bs="re")+s(HGV,by=Direction),data=Fires)
#summary(fit.gcv.re)
#gam.vcomp(fit.gcv.re)
#fit.gcv.re$aic
#plot(fit.gcv.re)
#plot(residuals(fit.gcv.re))
```

## Conlusion
I think that the best model is the Poisson model even though I am not sure it is very good. I am very bad at interpreting the residuals so if you are better at it it would help ! The Fires distribution clearly looks Poisson for me, so that's why I tried to fit a Poisson model.\
I don't know how to add the Tunnel and company to the model to make it be really interesting. \
I tried to fit things with gam like in the exercises sessions but again I am not an expert and I may have done useless things !



## References
On R commands\
https://bookdown.org/aurora_tumminello/statistics_lab/the-stock-market-data.html#data-exploration\

On modelling\
https://www.cetu.developpement-durable.gouv.fr/IMG/pdf/4-etude_pannes_accidents_et_incendies_pai_c._willmann_.pdf



