---
title: "Fires clean"
author: "candice"
date: "2022-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Packages, echo=FALSE, include = FALSE}
lapply(c("dplyr","chron","ggplot2","tidyr","questionr","survival","forcats",
         "data.table","table1","lubridate", "ggpubr","viridis","finalfit",
         "ggpubr", "ggthemes", "gridExtra", "tidyverse", "rstatix","ggsci",
         "wesanderson","kableExtra", "naniar","boot","scales","psych","ggcorrplot","ggradar", "caret", "gam"), 
       library, character.only=TRUE)

Sys.setlocale("LC_TIME", "English")
setwd("C:/Users/candi/Desktop/ETUDES/EPFL1A/semestre 1/regression methods/project")
load("C:/Users/candi/Desktop/ETUDES/EPFL1A/semestre 1/regression methods/project/Fires.RData")
```

## Introduction
This report describes an analysis of the database "Accidents", in which the data was simulated based on a real dataset of events in French motorway tunnels. Here, we focus on modelling the accidents.


## Exploration 
First, let's look at the data and try to fit a model to our data distribution. 

```{r Accidents, echo=FALSE}
p<-data.frame(rpois(10000,mean(Fires$Fires)))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=0.3)+
  ggtitle("poisson")
```

To fit a model, let's look at the correlations between the different variables and their distributions. This will enable us to understand better how they are possibly linked. 

```{r , echo=FALSE}
## Plot of covariates against each other to look at dependencies (slide 55)
#numerical variables 
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}


#numerical variables
pairs(~ Fires + Traffic + HGV + Slope + Limit + Length, data=Fires, upper.panel = panel.cor, diag.panel = panel.hist)
```
It looks like Fires are correlated with Traffic, HGV and Length. \
Length and HGV are correlated too so we should take this is account when we will try to fit the model.


## Model 
*Variable selection for linear model*
The first naive way to do it is to do a backward elimination on all the variables. But it gives a model based on companies and tunnel mainly, which is not a true model. Indeed, those effects are random. So, we do the backward elimination without those variables. 

```{r AIC and BIC, echo=FALSE}
fullmodel <- lm(Fires ~ log(Traffic) + HGV + Slope + Limit+ log(Length) + Direction + Urban + Type + Length + Traffic, data = Fires)

# Backward elimination
model.step.b <- step(fullmodel, direction = 'backward')

```
```{r regression, echo=FALSE}
summary(model.step.b)
par(mfrow=c(2,2)) 
plot(model.step.b)
```
The multiple R squared is not that good and the Slope coefficient is not significant. This might not be the best model at all. The distribution was close to a poisson, so let's try to fit poisson models.

*Poisson model*
First, as previously, I do a variable selection based on the AIC and BIC. 
```{r selection, echo =FALSE}
fullmodel <- glm(Fires ~ log(Traffic) + HGV + Slope + Limit+ log(Length) + Direction + Urban + Type + Length + Traffic, family="poisson", data=Fires)
model <- step(fullmodel, direction = 'backward')
``` 

The final model is the following one : 
```{r poisson model, echo =FALSE}
summary(model)

par(mfrow=c(2,2)) 
plot(model)

```
This model can probably be improved since some coefficients are not significant. Let's try to see if we can improve the model. 

```{r better model, echo=FALSE}
m2 <- glm(Fires ~ log(Traffic) + HGV + log(Length), family="poisson", data=Fires)
summary(m2)

par(mfrow=c(2,2)) 
plot(model)
``` 

*Random effects*
The idea here would be to consider random and fixed effects on the variables that were significant previously :\
  *random : Tunnel, Company, 
  *fixed : log(Traffic) + HGV + Urban + Type + log(Length) + Limit + SlopeType + Width\

First, we put Tunnel as the random effect
  
```{r Mixed model tunnel, echo=FALSE}
library('lme4')
summary(glmer(Fires ~ log(Traffic) + HGV + log(Length) + (1 | Tunnel ), data = Fires, family=poisson))
``` 
Then with company
```{r Mixed model company, echo=FALSE}
library('lme4')
summary(glmer(Fires ~ log(Traffic) + HGV + log(Length) + (1 | Company ), data = Fires, family=poisson))
``` 





