---
title: "Project Regression Methods - 2022"
author: "Céline Guex - sciper n°310673, Candice Baud - sciper n°359523"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
  includes:
    in_header: Preamble.tex
  pdf_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
lapply(c("dplyr","chron","ggplot2","tidyr","questionr","survival","forcats",
         "data.table","table1","lubridate", "ggpubr","viridis","finalfit",
         "ggpubr", "ggthemes", "gridExtra", "tidyverse", "rstatix","ggsci",
         "wesanderson","kableExtra",      "naniar", "boot", "scales", "psych",          "ggcorrplot", "ggradar", "caret", "gam"), 
       library, character.only=TRUE)
```

```{r loading-data}
load("Data/Breakdowns.RData")
load("Data/Accidents.RData")
load("Data/Fires.RData")
Fi<-Fires
Fi$Direction <- as.factor(Fi$Direction)
Fi$Urban <- as.factor(Fi$Urban)
Fi$Type <- as.factor(Fi$Type)
Fi$Limit <- as.factor(Fi$Limit)
Fi$SlopeType <- as.factor(Fi$SlopeType)
Fi$Tunnel <- as.factor(Fi$Tunnel)
Fi$Company <- as.factor(Fi$Company)
Ac<- Accidents

```

# Initial data analysis
## Description of Fires dataset
```{r exploratory, echo=FALSE}
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

par(mfrow=c(1,2)) 

p<-data.frame(rpois(10000,mean(Fires$Fires)))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=0.3)+
  ggtitle("poisson")

suppressWarnings(pairs(~ Fires + Traffic + HGV + Length + SlopeType + Type + Company, data=Fires, upper.panel = panel.cor, diag.panel = panel.hist))
```

## Description of Accidents dataset
```{r Accidents corr and fit, echo=FALSE}
#exponential dist
p<-data.frame(rexp(100,1/mean(Accidents$Acc)))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.5, lwd=1, bw=0.5)+
  ggtitle("exp(0.25)")


p<-data.frame(rpois(10000,mean(Accidents$Acc)))
ggplot(p,aes(x=p[,1]))+
  geom_histogram(aes(y=..density..), alpha=0.4, position="identity", lwd=0.1)+
  geom_density(alpha=0.3, lwd=1, bw=0.3)+
  ggtitle("Fitted poisson distribution")

#function below can be erased if we keep the one above for Fires
panel.hist <- function(x, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}


suppressWarnings(pairs(~ Acc + Traffic + Limit + Lanes + Urban + SlopeType ,  data=Accidents, upper.panel = panel.cor, diag.panel = panel.hist))

```
## Description of Breakdowns dataset

# Model fitting
## Fires
### Poisson model
```{r selection, echo =FALSE, include=FALSE}
library('glmtoolbox')
fullmodel <- glm(Fires ~ log(Traffic) + HGV + Slope + Limit+ log(Length) + Direction + Urban + Type + Length + Traffic, family="poisson", data=Fires)
model <- stepCriterion(fullmodel, direction = 'backward')
``` 

The best model according to AIC includes the following variables :
```{r selection result, echo =FALSE}
model$final
``` 
```{r better model, echo=FALSE}
m2 <- glm(Fires ~ log(Traffic) + HGV + log(Length) + Traffic, family="poisson", data=Fires)
summary(m2)
#par(mfrow=c(2,2)) 
#plot(m2)
```

The model we fitted implies this equation :
$$
\ln(\mathbb{E}(Fires)) = -27.70 + 1.178*\log(Traffic) +6.386*HGV + 0.6967*\log(Length) -1.19*10**9*Traffic
$$
```{r residuals model, echo=FALSE}
plot(m2$fitted.values, m2$residuals)
```

### Poisson mixed model
```{r Mixed model tunnel, echo=FALSE}
library('lme4')
mixed_model <- glmer(Fires ~ log(Traffic) + HGV + log(Length) + (1 | Tunnel ), data = Fires, family=poisson)
summary(mixed_model)
```

```{r Mixed model tunnel residuals, echo=FALSE}
library("lme4")
plot(mixed_model)
``` 

## Accidents
### Poisson model without the variables Tunnel and Company

```{r rescale variable accidents, echo=FALSE}
Accidents <- Accidents %>% mutate(
  Acc_sc = 0.5*sqrt(Accidents$Acc)
)

summary(Accidents$Acc_sc)
#We have that the mean is different than the standard deviation so we want to equalize them to apply the poisson model
#y <- Accidents$Acc
#y_prime <- 0.5*sqrt(y)

#mean(y_prime)
#sd(y_prime)
``` 


```{r variable selection, echo=FALSE}
library('glmtoolbox')
fullmodel <- glm(Acc_sc ~ Lanes + Direction + log(Traffic) + HGV + Slope + Urban + Type + log(Length) + Limit + SlopeType + Width + Lanes + Year, family=poisson, data=Accidents)
summary(fullmodel)

fullmodel2 <- glm(Acc ~ Lanes + Direction + log(Traffic) + HGV + Slope + Urban + Type + log(Length) + Limit + SlopeType + Width + Lanes + Year, family=poisson, data=Accidents)
summary(fullmodel2)


stepCriterion(fullmodel, criterion = 'aic', direction = c('forward', 'backward'))
``` 


```{r poisson model, echo=FALSE}
model <- glm(Acc ~ log(Traffic) + SlopeType + log(Length) + Limit + Type + Width + Direction + Slope + HGV + Year , family=poisson, data=Accidents)
summary(model)
``` 

The model equation we obtained is put below. 
$$
\ln(\mathbb{E}(Accidents)) = -51.286647 + 0.626838*\log(Traffic) + 0.582956*\log(Length) -0.018712*Limit -0.614479*Width -2.480020*Slope -0.867039*HGV + 0.020613 -0.139456*Direction2 + 0.947456*TypeUnidirectionnal +  0.886806*SlopeTypeBasin + 1.022039*SlopeTypeUnknown + 0.144215*SlopeTypeContinuous -0.590130*SlopeTypeRoof + 
$$

```{r poisson model residuals, echo=FALSE}
plot(model$fitted.values, model$residuals)
```

### Poisson model with random effects

```{r model selection test, echo=FALSE}
library("cAIC4")
mod1 <- lmer(Acc ~ log(Traffic) + log(Length) + (1|Company) + (1|Tunnel), data=Accidents)
stepcAIC(mod1, groupCandidates=NULL, fixEfCandidates = c('Slope','Type', 'Width', 'HGV'), data=Accidents, trace=TRUE, direction="both", keep = list('Company','Tunnel'))
```


```{r mixed model without rescaled variables, echo=FALSE}
library('lme4')
#log(Traffic) ok 5317
#log(Traffic) + log(Length) ok 5309
#log(Traffic) + log(Length) + Slope ok AIC 5297
# avec limit pb
#log(Traffic) + log(Length) + Slope + Type AIC 5294
#avec HGV fail to converge
#log(Traffic) + log(Length)  + HGV 5311
#avec Year fail 
#avec Direction fail 
#log(Traffic) + log(Length) + Slope + Type + Width 5291
#avec slopeType fail
modeltest<- glmer(Acc~ log(Traffic) + log(Length) + Slope + Type + Width + (1 | Tunnel) + (1 | Company), family=poisson, data = Accidents)
summary(modeltest)
``` 


```{r mixed model without rescaled variables residuals, echo=FALSE}
plot(modeltest)
```


## Conclusion on the dataset Accidents
If we consider models without random effects, the best Poisson model we can build in terms of AIC is the one with the following variables : log(Traffic), SlopeType, log(Length), Limit, Type, Width, Direction, Slope, HGV, Year. This model however has an AIC of 6065 which is quite enormous. It includes a lot of variables and doesn't take in account the variables Tunnel and Company. \
Without rescaling variables, one can fit a poisson mixed models with Tunnel and Company as random effects. This gives a smaller AIC which seems then to fit better our data. However, without rescaling, the model doesn't converge. \
Finally, by rescaling the variables, we obtain a smaller AIC because we obtain a smaller likelihood due to the rescaling and the model converges. \

## Breakdowns


# Discussion 



# References 



